Review: Adaptive Introgression Expanded the Genetic Base of Crops during post-Domestication Spread

### 2018-02-08 paper.tex
update rice section of paper.tex

The story of Asian rice (\emph{Oryza sativa}) domestication is still debated, with hybridization between wild and domesticated rice contributing to the complexity of this crop's history.
Certain genetic and archaeobotanical evidence point toward independent domestications of the two prominent varietal groups \emph{japonica} and \emph{indica} from the wild species \emph{Oryza rufipogon} (\emph{rufipogon} hereafter) in the Yangtze Basin of China and the Indian Ganges plain, respectively \cite{fuller2010consilience}.
Other studies support a single domestication occurring 8,200-13,500 BP in the Yangtze Basin, with later divergence of \emph{japonica} and \emph{indica} \cite{molina2011molecular}.
Huang and colleagues \cite{Huang2012} developed a genetic map of rice variation, which they used to measure genetic distance between wild and domesticated rices at and around domesication loci for various geographical locations, finding that \emph{japonica} was likely domesticated in the Pearl River area of Guangxi province, China (just south of the Yangtze Basin), and that \emph{indica} was likely the result of hybridization between \emph{japonica} and local \emph{rufipogon} populations in Southern and South-eastern Asia. %Also, Huang includes the following quote: "We noticed that many loci with strong signals of selection were nearly identical in both indica and japonica where FST between indica and japonica was extremely low, indicating that introduction of traits during domestication has in many cases involved introgression events."

@@

Vaughan and colleagues \cite{vaughan2008evolving} posit a hypothesis combining elements of both theories.
According to this hypothesis, the non-shattering allele arose in a single cultivated rice population and then introgressed into many diverse cultivated populations (some japonica-like, some indica-like) in the region.
This would have facilitated the subsequent spread of other domestication alleles among those populations.
As these domesticated populations then spread into new environments, they would have recieved introgressions from locally-adapted wild relatives, retaining alleles that improved fitness (in agreement with \cite{second1982origin}).
Contemporary populations of wild and domesticated rice show evidence of this history of local adaptation.
The wild relatives \emph{rufipogon} and \emph{nivara} both maintain high genotypic and phenotypic diversity and exhibit locally-adaptive traits (\emph{rufipogon} is adapted to forested wetland environments, whereas \emph{nivara} is adapted to dryer conditions and has life cycle adaptations to survive grazing pressure), suggesting that either offered (and offers) adaptive genetic potential.
Likewise, domesticated rice varieties display patterns of local adaptation (for examples, two of the domesticated rice deepwater varieties (\emph{rayada} and \emph{ashwina}) are said to be selected for the environment along the Ganges river, the \emph{japonicas} are split into temperate and tropical subgroups, and the \emph{indicas} are best suited for lowland environments.

%Other studies support a single center of domestication, arguing that present-day patterns of rice variation could be explained by a single domestication event in a population with high standing variation, followed by dispersal into sympatry with locally-adapted wild relatives in diverse environments, genetic admixture, and selection for adaptive alleles \cite{vaughan2008evolving}.
%The high genetic diversity within domesticated rice is likely due to introgression from wild relatives both within the domestication center(s) and in new environments where rice has dispersed \cite{second1982origin}, and gene flow between domesticated Asian rice and its wild relatives outside of the Yangtze Basin could have imparted locally adaptive traits \cite{vaughan2008evolving}.
%The wild relatives \emph{rufipogon} and \emph{nivara} both maintain high genotypic and phenotypic diversity and exhibit locally-adaptive traits (whereas \emph{rufipogon} is adapted to forested wetland environments, \emph{nivara} is adapted to dryer conditions and has life cycle adaptations to survive grazing pressure).
%From Huang2012, "Based on the SNP data, the sequence diversity (p) of O. rufipogon was estimated at 0.003, which is higher than that in O. sativa—the sequence diversity is 0.0024 for O. sativa, and 0.0016 and 0.0006 for indica and japonica, respectively21".
%Likewise, domesticated rice varieties display patterns of local adaptation (for examples, two of the domesticated rice deepwater varieties (\emph{rayada} and \emph{ashwina}) are said to be selected for the environment along the Ganges river, the \emph{japonicas} are split into temperate and tropical subgroups, and the \emph{indicas} are best suited for lowland environments.
%Gene flow between wild and cultivated rices, though asymmetrical, is frequent enough that nuisance weedy hybrids commonly arise amongst cultivated populations and must be managed with difficulty \cite{vaughan2008evolving}.

To date, research into adaptive introgression in the domestication of rice has been insufficient to detect clearly-supported examples (though \cite{zhao2010genomic} emphasizes the necessity of this research).
%\cite{zhao2010genomic} used a SNP panel and STRUCTURE analysis to uncover patterns of population structure, admixture, and introgression within domesticated rices, and the authors emphasize the importance of similar research that includes wild rice accessions.
There are perhaps some practical reasons why research has not yet been devoted to this inquiry.
As with many other domesticated crops, gene flow between wild and domesticated rices is highly asymmetric (estimates of wild rice admixture in domesticated rice are less than 5 percent \cite{wang2017asian}).
This asymmetry is due in part to the closed floret architecture of the domesticated rice, which hinders outcrossing.
Crop-wild hybridization is frequent enough that nuisance weedy hybrids commonly arise amongst cultivated populations and must be managed with difficulty \cite{vaughan2008evolving}.
During early domestication, introgression may have been more prevalent than at present because barriers to crop-to-wild introgression may have been less severe and because the inbreeding reproductive system of rice would not have been as firmly established \cite{vaughan2008evolving}.
Furthermore, the contemporary distribution of wild rice does not capture the range and diversity of wild rice during early domestication and range expansion of rice.

%Introgression into indica is expected to be more likely than japonica, due to the higher degree of sympatry between indica and its wild relatives.

%note from Asian wild rice is a hybrid swarm with extensive gene flow and feralization from domesticated rice:
%Cite this paper to talk about how the ranges of the several subspecies of domesticated rice are distinct/sometimes overlap.  The paper makes the point that this overlap helps facilitate gene flow.

>>>>

The story of Asian rice (\emph{Oryza sativa}) domestication is still debated.
Certain genetic and archaeobotanical evidence points toward independent domestications of the two prominent varietal groups \emph{japonica} and \emph{indica} from the wild species \emph{Oryza rufipogon} (\emph{rufipogon} hereafter) in the Yangtze Basin of China 7000 BP and the Indian Ganges plain 4000 BP, respectively \cite{fuller2010consilience}, with possibly another domestication event giving rise to \emph{aus} in Bangladesh or central India \cite{civavn2015three}.
Other studies support a single domestication occurring 8,200-13,500 BP in the Yangtze Basin, with later divergence of \emph{japonica} and \emph{indica} \cite{molina2011molecular, Huang2012}.
Huang and colleagues \cite{Huang2012} developed a genetic map of rice variation, which they used to measure genetic distance between wild and domesticated rices at and around domesication loci for various geographical locations, finding that \emph{japonica} was likely domesticated in the Pearl River area of Guangxi province, China (just south of the Yangtze Basin), and that \emph{indica} was likely the result of hybridization between \emph{japonica} and local \emph{rufipogon} populations in Southern and South-eastern Asia.
In a re-examination of these same data, \cite{civavn2015three} found evidence supporting independent domestications of \emph{japonica}, \emph{indica}, and \emph{aus}, as well as a hybrid origin (\emph{japonica} x \emph{aus}) of variety \emph{aromatic}.
A third analysis \cite{choi2018multiple} compared these two disparate results and the methods utilized therein, concluding that domestication alleles (including \emph{LABA1}, \emph{PROG1}, and \emph{sh1}) arose via a single domestication event of \emph{japonica}, and that these alleles introgressed into several wild \emph{rufipogon} subpopulations which became the progenitors to the other Asian rice varieties.
 %Also, Huang includes the following quote: "We noticed that many loci with strong signals of selection were nearly identical in both indica and japonica where FST between indica and japonica was extremely low, indicating that introduction of traits during domestication has in many cases involved introgression events."

@@

The findings of \cite{choi2018multiple} run in parallel with a hypothesis posited by Vaughan and colleagues in 2008 \cite{vaughan2008evolving}.
%Vaughan and colleagues \cite{vaughan2008evolving} posit a hypothesis combining elements of both theories.
According to this hypothesis, the non-shattering allele (\emph{sh1}) arose in a single cultivated rice population and then introgressed into many diverse cultivated populations (some \emph{japonica}-like, some \emph{indica}-like) in the region.
This would have facilitated the subsequent spread of other domestication alleles among those populations.
As these domesticated populations then spread into new environments, they would have received introgressions from locally-adapted wild relatives, retaining alleles that improved fitness (in agreement with \cite{second1982origin}).
Both Vaughan and Choi support the role of introgression between domesticated Asian rice and locally-adapted wild \emph{rufipogon} populations in the development and dispersal of Asian rice into new environments.

Despite many conditions favorable to their prevalence, to date, individual examples of adaptive alleles naturally introgressing into domesticated Asian rice are not clear (though \cite{zhao2010genomic} emphasizes the necessity of research on this topic).
Contemporary populations of wild rices show evidence of this history of local adaptation.
Wild \emph{rufipogon} and \emph{nivara} both maintain high genotypic and phenotypic diversity and exhibit locally-adaptive traits (\emph{rufipogon} is adapted to forested wetland environments, whereas \emph{nivara} is adapted to dryer conditions and has life cycle adaptations to survive grazing pressure), suggesting that either offered (and offers) adaptive genetic potential.
Domesticated varieties have likewise adapted to local environments.
The \emph{japonicas} are generally more adapted to upland temperate regions (and are split into subgroups adapted to more temperate or more tropical conditions), the \emph{indicas} are best suited for tropical lowland environments, \emph{aus} matures quickly and tolerates drought stress, and two deepwater varieties (\emph{rayada} and \emph{ashwina}) are said to be selected for the environment along the Ganges river \cite{vaughan2008evolving, garris2005genetic}.
Crop-wild hybridization is frequent enough that nuisance weedy hybrids commonly arise amongst cultivated populations and must be managed with difficulty \cite{vaughan2008evolving}.
However, as with many other domesticated crops, gene flow between wild and domesticated rices is highly asymmetric (estimates of wild rice admixture in domesticated rice are less than 5 percent \cite{wang2017asian}).
This asymmetry is due in part to the closed floret architecture of the domesticated rice, which hinders outcrossing.
Still, barriers to crop-to-wild introgression may have been less severe during early domestication as the inbreeding reproductive system of domesticated rice would not have been as firmly established \cite{vaughan2008evolving}.
Furthermore, the contemporary distribution of wild rice does not capture the full range and diversity of wild rice during early domestication and range expansion of Asian rice.
